Efficient entropy stable Gauss collocation methods

### 2018-08-05 docs/ESDG_GQ/esdg_gq.tex
updating ESDG with EC shock vortex

This paper focuses on the construction of efficient semi-discretely entropy stable collocation schemes based on Gauss nodes, using the framework introduced in \cite{chan2017discretely, chan2018discretely}.  While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce significantly more communication and computation than collocation points which contain boundary nodes.  This work introduces new inter-element coupling terms which avoid this all-to-all coupling, requiring only communication of face values between neighboring elements.

@@

(see Figure~\ref{fig:twoelem})
\note{Add pics of two-element mesh}

@@

\section{Extension to higher dimensions and curved meshes}
\label{sec:2}

@@

Let the domain now be decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of $\hat{D}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.  We define geometric terms $\bm{G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$
\[
J^k\bm{G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_j},
\]
where $J^k$ is the determinant of the Jacobian of $\bm{\Phi}^k$.  We also introduce the face Jacobian $J^k_f$, which is computed from the mapping of a physical face to a reference face.  We define matrices over each physical element $D^k$ 
\begin{gather*}
\bm{W}_k = \bm{W} \diag{\bm{J}^k}, \qquad \bm{B}^i_k = \bm{B}^i_N \diag{\bm{n}_i \circ\bm{J}^f }\\
\bm{Q}^i_k = \frac{1}{2} \sum_{j=1}^d \LRp{\diag{\tilde{\bm{G}}^k_{ij}}\bm{Q}^i_N  + \bm{Q}^i_N \diag{\tilde{\bm{G}}^k_{ij}} },
\end{gather*}
where we have discretized the differentiation matrix in split form \cite{nordstrom2006conservative}.  
Here, $\bm{J}^k$ denotes the vector of values of $J^k$ at volume quadrature points, $\tilde{\bm{G}}^k_{ij}$ denotes the vector of values of $J^k\bm{G}^k_{ij}$ at volume and face quadrature points, and $\bm{n}^k_i, \bm{J}^k_f$ denote the $i$th outward unit normal and $J^k_f$ at face quadrature points.  Then, an entropy conservative scheme in two dimensions is given as follows: 
\begin{theorem}
\label{thm:esdg2d}
Assume that $\bm{Q}^k\bm{1} = 0$.  Let $\bm{u}^k_N$ denote the solution on $D^k$ satisfying
\begin{gather*}
\sum_{k}\LRp{\bm{W}_k\td{\bm{u}^k_N}{t} + \sum_{i=1}^d \LRp{2\LRp{\bm{Q}^i_k\circ\bm{F}_S} + \bm{V}_f^T\bm{B}_k^i\bm{f}_S\LRp{{\tilde{\bm{u}}_f}^+,\tilde{\bm{u}}^k_f}}} = 0\\
\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \begin{bmatrix}
\bm{u}^k_N\\
\tilde{\bm{u}}^k_f
\end{bmatrix}, \qquad \tilde{\bm{u}}^k_f = \bm{u}\LRp{\bm{v}^k_f}, \qquad 
\bm{v}^k_f = {\bm{V}_f\bm{v}\LRp{\bm{u}^k_N}},
\end{gather*}
where $\tilde{\bm{u}}^+$ denotes boundary values of the entropy-projected conservative variables on neighboring elements.  Then, $\bm{u}^k_N$ satisfies the discrete conservation of entropy 
\[
\sum_{k} \LRp{\bm{1}^T\bm{W}\td{\bm{u}^k_N}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi\LRp{\tilde{\bm{u}}_f}}} = 0.
\]
\end{theorem}
\begin{proof}
The proof is a special case of the proof of Theorem 1 in \cite{chan2018discretely}.  
\end{proof}

We note that Theorem~\ref{thm:esdg2d} relies on the assumption that $\bm{Q}^k\bm{1} = 0$.  This condition is equivalent to ensuring that the scaled geometric terms $\tilde{\bm{G}}^k_{ij}$ satisfy a discrete geometric conservation law (GCL) \cite{carpenter2014entropy, gassner2017br1, crean2018entropy, chan2018discretely}.  For two-dimensional degree $N$ (isoparametric) mappings, the GCL is automatically satisfied.  However, in three dimensions, the GCL is not guaranteed to be satisfied at the discrete level, due to the fact that geometric terms for isoparametric mappings are polynomials of degree higher than $N$.

@@

For $N_{\rm geo} \geq \left\lfloor\frac{N}{2}\right\rfloor + 1$, modifications to the computation of geometric terms are required to ensure that the GCL is satisfied at the discrete level.  For general SBP operators, the discrete GCL can be enforced through the solution of a local least squares problem \cite{crean2018entropy}.  We take an alternative approach and  construct geometric terms using the approach of Kopriva \cite{kopriva2006metric}.  This construction takes advantage of the fact that GLL and Gauss collocation methods correspond to polynomial discretizations.  Kopriva's construction is based on rewriting the geometric terms as the curl of some auxiliary quantity.  By approximating this auxiliary quantity using polynomial interpolation prior to applying the curl, the geometric terms are approximated by degree $N$ polynomials which satisfy the discrete GCL by construction.  %We note that this approach can also be adapted to more general geometric mappings (e.g.\ rational or superparametric).  These mappings can offer additional geometric accuracy, which can be advantageous for problems with wall boundary conditions \cite{toulorge2016optimizing, zwanenburg2017necessity}.

To summarize, entropy stable Gauss collocation schemes on curved meshes require the following additional steps:
\begin{enumerate}
\item Construct polynomial approximations of the geometric terms at GLL nodes using (\ref{eq:curlcons}) \cite{kopriva2006metric}.  
\item Evaluate the geometric terms at volume and surface quadrature points.  
\item Compute normal vectors using (\ref{eq:normal}).  
\end{enumerate}

@@

p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \\

@@

@@

\begin{figure}
\centering
\subfloat[Affine mesh]{
\begin{tikzpicture}
\begin{loglogaxis}[
    width=.5\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ errors}, 
%    xmin=.0125, 
    ymin=1e-6, ymax=5,
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
    legend entries={GLL, Gauss}    
]
\pgfplotsset{
cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
}
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,2.62759)(0.125,0.728523)(0.0625,0.112225)(0.03125,0.018711)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,1.53203)(0.125,0.23283)(0.0625,0.0417033)(0.03125,0.00806707)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.934758)(0.125,0.191949)(0.0625,0.0119788)(0.03125,0.000558681)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.562488)(0.125,0.0864096)(0.0625,0.00268061)(0.03125,0.000136725)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.460425)(0.125,0.02445)(0.0625,0.000821249)(0.03125,4.67491e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.239027)(0.125,0.0077433)(0.0625,0.000219705)(0.03125,2.67589e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.138105)(0.125,0.00744055)(0.0625,0.000126057)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.0547127)(0.125,0.00242675)(0.0625,4.2962e-05)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.0489719)(0.125,0.000897467)(0.0625,2.55644e-05)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.0246519)(0.125,0.000324802)(0.0625,2.44009e-05)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.023319)(0.125,0.00021951)(0.0625,NaN)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.00801774)(0.125,7.39203e-05)(0.0625,NaN)(0.03125,NaN)};

@@

    ymin=1e-6, ymax=5,
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	

@@

\note{describe shock-vortex problem}

We modify the shock-vortex interaction problem such that wall boundary conditions are imposed at top and bottom boundaries.  For left and right boundaries, we impose periodic boundary conditions.

Experiments using the matrix dissipation flux \cite{winters2017uniquely}.

We note that entropy stable wall boundary conditions can also be imposed when using decoupled SBP operators with over-integration as well \cite{chan2017discretely}.  As far as the authors know, the stable and high order accurate imposition of such boundary conditions for existing GSBP couplings \cite{crean2017high} remains an open problem.

@@

>>>>

While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce an ``all-to-all'' coupling between degrees of freedom on two neighboring elements.  These coupling terms require evaluating two-point fluxes between solution states at all collocation nodes on two neighboring elements, resulting in significantly more communication and computation compared to collocation schemes based on point sets containing boundary nodes.  This work introduces efficient and entropy stable inter-element coupling terms for Gauss collocation schemes which require only communication of face values between neighboring elements.  The construction of these terms follows the framework introduced in \cite{chan2017discretely, chan2018discretely} for triangles and tetrahedra.

The main motivation for exploring tensor product (quadrilateral and hexahedral) elements is the significant reduction in the number of operations required compared to high order entropy stable schemes on simplicial meshes \cite{chan2017discretely, chan2018discretely}.  Entropy stability schemes on simplicial elements require evaluating two-point fluxes between solution states at all quadrature points on an element.  The number of quadrature points on a simplex scales as $O(N^d)$ in $d$ dimensions, and results in $O(N^{2d})$ two-point flux evaluations per element.  In contrast, entropy stable schemes on quadrilateral and hexahedral elements require only the evaluation of two-point fluxes along lines of nodes due to a tensor product structure, resulting in $O(N^{d+1})$ evaluations in $d$ dimensions.

@@

%(see Figure~\ref{fig:twoelem})
%\note{Add pics of two-element mesh}

@@

\section{Extension to higher dimensions and non-affine meshes}
\label{sec:2}

@@

Let the domain now be decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of $\hat{D}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.  We define geometric terms ${G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$
\[
{G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_j},
\]
where $J^k$ is the determinant of the Jacobian of $\bm{\Phi}^k$.  We also introduce the face Jacobian $J^k_f$, which is computed from the mapping of a physical face to a reference face.  We define matrices over each physical element $D^k$ 
\begin{gather}
\bm{W}_k = \bm{W} \diag{\bm{J}^k}, \qquad \bm{B}^i_k = \bm{B}^i_N \diag{\bm{n}_i \circ\bm{J}^f } \nonumber\\
\bm{Q}^i_k = \frac{1}{2} \sum_{j=1}^d \LRp{\diag{{\bm{G}}^k_{ij}}\bm{Q}^i_N  + \bm{Q}^i_N \diag{{\bm{G}}^k_{ij}} } \label{eq:splitcurv},
\end{gather}
where we have discretized the curved differentiation matrix in split form \cite{nordstrom2006conservative}.  
Here, $\bm{J}^k$ denotes the vector of values of $J^k$ at volume quadrature points, ${\bm{G}}^k_{ij}$ denotes the vector of values of ${G}^k_{ij}$ at volume and face quadrature points, and $\bm{n}^k_i, \bm{J}^k_f$ denote the $i$th outward unit normal and $J^k_f$ at face quadrature points.  Then, an entropy conservative scheme in two dimensions is given as follows: 
\begin{theorem}
\label{thm:esdg2d}
Assume that $\bm{Q}^k\bm{1} = 0$, and that $D^k$ are mapped curvilinear elements.  Let $\bm{u}^k_N$ denote the solution on $D^k$ satisfying
\begin{gather*}
\sum_{k}\LRp{\bm{W}_k\td{\bm{u}^k_N}{t} + \sum_{i=1}^d \LRp{2\LRp{\bm{Q}^i_k\circ\bm{F}_S} + \bm{V}_f^T\bm{B}_k^i\bm{f}_S\LRp{{\tilde{\bm{u}}_f}^+,\tilde{\bm{u}}^k_f}}} = 0\\
\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \begin{bmatrix}
\bm{u}^k_N\\
\tilde{\bm{u}}^k_f
\end{bmatrix}, \qquad \tilde{\bm{u}}^k_f = \bm{u}\LRp{\bm{v}^k_f}, \qquad 
\bm{v}^k_f = {\bm{V}_f\bm{v}\LRp{\bm{u}^k_N}},
\end{gather*}
where $\tilde{\bm{u}}^+$ denotes boundary values of the entropy-projected conservative variables on neighboring elements.  Then, $\bm{u}^k_N$ satisfies the discrete conservation of entropy 
\[
\sum_{k} \LRp{\bm{1}^T\bm{W}\td{\bm{u}^k_N}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi\LRp{\tilde{\bm{u}}_f}}} = 0.
\]
\end{theorem}
The proof is a special case of the proof of Theorem 1 in \cite{chan2018discretely}, with the quadrature rule taken to be a tensor product rule with $(N+1)$ Gauss (or GLL) points in each coordinate direction.

\begin{remark}
The operator $\bm{Q}^i_k$  in (\ref{eq:splitcurv}) can be applied without needing to explicitly store geometric terms $G^k_{ij}$ at face quadrature points.  By using the structure of the boundary matrix $\bm{B}^i$, expressions involving geometric terms on faces can be replaced by expressions involving components of unit outward normals $\bm{n}_i$ on $D^k$.  
\end{remark}

We note that Theorem~\ref{thm:esdg2d} relies on the assumption that $\bm{Q}^k\bm{1} = 0$.  This condition is equivalent to ensuring that the scaled geometric terms ${\bm{G}}^k_{ij}$ satisfy a discrete geometric conservation law (GCL) \cite{carpenter2014entropy, gassner2017br1, crean2018entropy, chan2018discretely}.  For two-dimensional degree $N$ (isoparametric) mappings, the GCL is automatically satisfied.  However, in three dimensions, the GCL is not guaranteed to be satisfied at the discrete level, due to the fact that geometric terms for isoparametric mappings are polynomials of degree higher than $N$.

@@

For $N_{\rm geo} \geq \left\lfloor\frac{N}{2}\right\rfloor + 1$, modifications to the computation of geometric terms are required to ensure that the GCL is satisfied at the discrete level.  For general SBP operators, the discrete GCL can be enforced through the solution of a local least squares problem \cite{crean2018entropy}.  We take an alternative approach and  construct geometric terms using the approach of Kopriva \cite{kopriva2006metric}.  This construction takes advantage of the fact that GLL and Gauss collocation methods correspond to polynomial discretizations.  Kopriva's construction is based on rewriting the geometric terms as the reference curl of an interpolated auxiliary quantity
\begin{equation}
\label{eq:conscurl}
\begin{bmatrix}
\bm{G}^k_{1j}\\
\bm{G}^k_{2j}\\
\bm{G}^k_{3j}
\end{bmatrix} = \begin{bmatrix}
\LRp{-\hat{\Grad} \times I_N\LRp{z\hat{\Grad}y}}_j\\
\LRp{\hat{\Grad} \times I_N\LRp{z\hat{\Grad}x}}_j\\
\LRp{\hat{\Grad} \times I_N\LRp{x\hat{\Grad}y}}_j
\end{bmatrix}, \qquad j = 1,2,3.
\end{equation}
Here, $I_N$ denotes the polynomial interpolation operator using GLL nodes.  By interpolating the auxiliary quantity in (\ref{eq:conscurl}) using polynomial interpolation prior to applying the curl, the geometric terms are approximated by degree $N$ polynomials which satisfy the discrete GCL by construction.  These GCL-satisfying geometric terms can then be used to compute normal vectors.  For a watertight mesh, the constructed normal vectors are guaranteed to be continuous across faces \cite{chan2018discretely}.  %We note that this approach can also be adapted to more general geometric mappings (e.g.\ rational or superparametric).  These mappings can offer additional geometric accuracy, which can be advantageous for problems with wall boundary conditions \cite{toulorge2016optimizing, zwanenburg2017necessity}.

To summarize, extending higher dimensional entropy stable Gauss collocation schemes to curved meshes requires the following steps: 
\begin{enumerate}
\item Construct polynomial approximations of the geometric terms using equation (\ref{eq:conscurl}) and interpolation at GLL nodes \cite{kopriva2006metric}.  
\item Evaluate approximate geometric terms at volume and surface points, and compute normal vectors in terms of the approximate geometric terms.  
\item Compute physical derivatives using the split form (\ref{eq:splitcurv}).  
\end{enumerate}
Apart from evaluating the GCL-satisfying geometric terms at separate volume and surface points, entropy stable Gauss collocation schemes are extended to curved meshes in the same manner as GLL and staggered-grid collocation schemes \cite{carpenter2014entropy, parsani2016entropy, chan2018discretely}.

@@

\begin{gather}
p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \label{eq:fluxaux} \\

@@

\end{gather}

@@

\begin{figure}
\centering
\subfloat[Affine mesh]{
\begin{tikzpicture}
\begin{loglogaxis}[
    width=.5\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ errors}, 
%    xmin=.0125, 
    ymin=1e-5, ymax=5,
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
    legend entries={GLL, Gauss}    
]
\pgfplotsset{
cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
}
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,2.62759)(0.125,0.728523)(0.0625,0.112225)(0.03125,0.018711)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,1.53203)(0.125,0.23283)(0.0625,0.0417033)(0.03125,0.00806707)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.934758)(0.125,0.191949)(0.0625,0.0119788)(0.03125,0.000558681)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.562488)(0.125,0.0864096)(0.0625,0.00268061)(0.03125,0.000136725)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.460425)(0.125,0.02445)(0.0625,0.000821249)(0.03125,4.67491e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.239027)(0.125,0.0077433)(0.0625,0.000219705)(0.03125,2.67589e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.138105)(0.125,0.00744055)(0.0625,0.000126057)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.0547127)(0.125,0.00242675)(0.0625,4.2962e-05)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.0489719)(0.125,0.000897467)(0.0625,2.55644e-05)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.0246519)(0.125,0.000324802)(0.0625,2.44009e-05)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.023319)(0.125,0.00021951)(0.0625,NaN)(0.03125,NaN)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.25,0.00801774)(0.125,7.39203e-05)(0.0625,NaN)(0.03125,NaN)};

@@

    ymin=1e-5, ymax=5,
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	

@@

The next problem considered is the shock-vortex interaction described in \cite{shu1998essentially}.  The domain is taken to be $[0,2]\times [0,1]$, and is triangulated with uniform quadrilateral elements.  Wall boundary conditions are imposed on the top and bottom boundaries, and inflow boundary conditions are typically imposed on the left and right boundaries.  We modify the problem setup such that periodic boundary conditions are imposed at the left and right boundaries.  Wall boundary conditions are imposed using a mirror state for the normal velocity, which was shown to be entropy stable in \cite{svard2014entropy, chen2017entropy}.

The initial condition is taken to be the superposition of a stationary shock and a vortex propagating towards the right.  The stationary Mach $M_s = 1.1$ shock is positioned at $x = .5$ normal to the $x$ axis, with left state $\LRp{\rho_L, u_L, v_L, p_L} = \LRp{1, \sqrt{\gamma}, 0, 1}$.  The right state is a scaling of the left state computed using the Rankine-Hugoniot conditions, such that the ratio of upstream and downstream states is
%$\LRp{\rho_R, u_R, v_R, p_R} = \LRp{\rho_L/ \alpha_\rho, u_L / \alpha_u, 0, p_L/\alpha_p}$, where
\begin{align*}
\frac{\rho_L}{\rho_R} = \frac{u_L}{u_R} = \frac{2+ (\gamma-1) M_s^2}{(\gamma+1)M_s^2}, \qquad \frac{p_L}{p_R} = 1+ \frac{2\gamma}{\gamma+1}\LRp{M_s^2-1}, \qquad v_R = 0.
\end{align*}
The isentropic vortex is centered at $(x_c,y_c) = (.25, .5)$ and given in terms of velocity fluctuations $\delta u$ and $\delta v$, which are functions of the tangential velocity $v_{\theta}$
\[
\delta u = v_{\theta} \sin(\theta), \qquad
\delta v = -v_{\theta} \cos(\theta), \qquad
v_{\theta} = \epsilon \tau e^{\alpha(1-\tau^2)},
%u = \epsilon \tau e^{\alpha(1-\tau^2)}\sin(\theta), \qquad
%v = -\epsilon \tau e^{\alpha(1-\tau^2)}\cos(\theta), \qquad
%\delta T = -\frac{(\gamma-1)\epsilon^2e^{2\alpha(1-\tau^2)}}{4\alpha\gamma}.
\]
where $r = \sqrt{(x-x_c)^2 + (y-y_c)^2}$ is the radius from the vortex center, $\tau = r/r_c$, and $\theta = \tan^{-1}\LRp{\frac{y-y_c}{x-x_c}}$.  We follow \cite{shu1998essentially} and take $\epsilon = .3$, $\alpha = .204$, and $r_c = .05$.  The vortex temperature is computed as a fluctuation $\delta T$ of the upstream state $T_L = p_L/\rho_L$
\[
\delta T =  - \frac{(\gamma-1)\epsilon^2 e^{2\alpha (1-\tau^2)}}{4\alpha\gamma}.  
\]
The vortex density and pressure are computed using an isentropic assumption.  To summarize, the initial condition for the shock-vortex interaction problem is 
\begin{gather*}
\rho = \rho_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}}, \qquad u = u_s + \delta u, \qquad v = v_s + \delta v, \qquad p = p_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}},
\end{gather*}
where $(\rho_s,  u_s, v_s, p_s)$ denote the discontinuous stationary shock solution given by the left and right states $\LRp{\rho_L, u_L, v_L, p_L}, \LRp{\rho_R, u_R, v_R, p_R}$.

\begin{figure}
\centering
\subfloat[Entropy conservative flux, $T = .3$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp3_EC.png}}
\hspace{.05em}
\subfloat[Entropy conservative flux, $T = .7$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp7_EC.png}}\\
\subfloat[Lax-Friedrichs flux, $T = .3$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp3_LF.png}}
\hspace{.05em}
\subfloat[Lax-Friedrichs flux, $T = .7$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp7_LF.png}}\\
\subfloat[Matrix dissipation flux, $T = .3$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp3.png}}
\hspace{.05em}
\subfloat[Matrix dissipation flux, $T = .7$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp7.png}}
\caption{Shock vortex solution at time $T = .7$ using entropy stable Gauss collocation schemes with $N=4, h = 1/100$.}
\label{fig:shockvort}
\end{figure}

We compare three different entropy stable Gauss collocation schemes.  All three utilize the entropy conservative flux of Chandrashekar \cite{chandrashekar2013kinetic}.  For the first scheme, we do not introduce any additional interface dissipation, which produces an entropy conservative scheme.  The second scheme introduces an entropy-dissipative interface term using Lax-Friedrichs penalization, while the third scheme utilizes the matrix dissipation flux introduced in \cite{winters2017uniquely}.  This flux adds a dissipation of the form $\bm{R}\bm{D}\bm{R}^T\jump{{\bm{v}}^k_f}$, where $\jump{{\bm{v}}^k_f}$ denotes the jump in the entropy variables.  In two dimensions, the matrices $\bm{R}, \bm{D}$ are 
\begin{align*}
\bm{R} &= \begin{bmatrix}
1 & 1 & 0 & 1\\
\avg{u}-\bar{a} n_x & \avg{u} & n_y & \avg{u}+\bar{a} n_x\\
\avg{v}-\bar{a} n_y & \avg{v} & -n_x & \avg{v}+\bar{a} n_y\\
h-\bar{a} \bar{u}_n & \frac{1}{2}\nor{\bm{u}}^2_{\rm avg} & \bar{u}_n n_y - \avg{v}n_x & h + \bar{a} \bar{u}_n
\end{bmatrix}\\
\bm{D} &= {\rm diag}{\begin{pmatrix}
\LRb{\bar{u}_n - a}\frac{\avg{\rho}^{\log}}{2\gamma}, & \LRb{\bar{u}_n}\frac{\avg{\rho}^{\log}(\gamma-1)}{\gamma}, & \LRb{\bar{u}_n}p_{\rm avg}, & \LRb{\bar{u}_n + a}\frac{\avg{\rho}^{\log}}{2\gamma} \end{pmatrix}},
\end{align*}
where $\bar{a}, \bar{u}_n$ are defined as
\begin{gather*}
\bar{u}_n = \avg{u}n_x + \avg{v}n_y, \qquad \bar{a} = \sqrt{\frac{\gamma p_{\rm avg}}{\avg{\rho}^{\log}}}, \qquad h = \frac{\gamma}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\bar{u}, 
\end{gather*}
and $p_{\rm avg}, \nor{\bm{u}}^2_{\rm avg}$ are defined as in (\ref{eq:fluxaux}).

%It is known that the Lax-Friedrichs flux is fairly dissipative compared to more nuanced fluxes \cite{harten1983upstream}, though this dissipation is less pronounced at higher polynomial degrees in practice.  
Figure~\ref{fig:shockvort} shows solutions for with $N=4$ and $h = 1/100$ Gauss collocation schemes using a non-dissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.  In all cases, the vortex passes through the shock stably without the use of additional slope limiting, filtering, or artificial viscosity.  However, the entropy conservative scheme produces a large number of spurious oscillations in the solution.  These are reduced away from the shock under Lax-Friedrichs dissipation, though oscillations still persist around a large neighborhood of the discontinuity.  The Gibbs-type oscillations are most localized under the matrix dissipation flux.

We note that this experiment also verifies that entropy stable decoupled SBP schemes (including the over-integrated case \cite{chan2017discretely}) are compatible with entropy stable wall boundary conditions.  As far as the authors know, the stable and high order accurate imposition of such boundary conditions for existing GSBP couplings described in \cite{crean2017high} and Section~\ref{sec:gsbpsat} remains an open problem.

@@

\begin{figure}
\centering
\subfloat[GLL collocation]{
\begin{tikzpicture}
\begin{axis}[
    width=.525\textwidth,
    xlabel={Time $t$},
    ylabel={$-\pd{\kappa}{t}$}, 
%    xmin=.0125, xmax=.75,
    ymin=-.001, ymax=.016,
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
    legend entries={Affine, Curved}    
]
\pgfplotsset{
cycle list={{blue}, {red, dashed}}
}
